---
title: "BART Parameter Recovery — STL, 4PAR, EWMV"
author: "Zach Tefertiller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(kableExtra); library(here)
})
source(here("analysis", "plotting", "recovery_plot.R"))
```

# Overview

Each of the three models is assessed the same way, in each of the three task
phases (blue = pre-reversal, orange = post-reversal, pink = control):

1. **True parameters** are drawn from that model × condition's population
   distribution — either the fitted group mean/SD if the fit objects are present,
   or a moment-match to the published per-participant posterior means if not.
   Drawing from a population rather than reusing the per-subject posterior means
   as "truth" keeps a realistic between-subject spread, so recovery is not
   deflated by shrinkage.
2. **Data are simulated** from the model's generative process on that condition's
   fixed breakpoints, taken from `data_published/bart_rl_online_trials.csv`.
3. **The model is refit** to the simulated data, and the true parameters are
   correlated with the recovered per-subject posterior means.

The simulators reproduce their `.stan` decision processes exactly, including the
numerical guards in 4PAR:

| Model | Simulator | Stan model |
|---|---|---|
| STL | `modeling/stl/simulate_stl.R` | `modeling/stl/stl_vec.stan` |
| 4PAR | `modeling/fourpar/simulate_fourpar.R` | `modeling/fourpar/4par.stan` |
| EWMV | `modeling/ewmv/simulate_ewmv.R` | `modeling/ewmv/ewmv_vec.stan` |

To generate the inputs for this report:

```r
Rscript modeling/run_recovery.R           # all 9 model x condition cells
Rscript modeling/run_recovery.R ewmv      # one model, all three conditions
Rscript modeling/run_recovery.R stl blue  # a single cell
```

Results land in `mcmc/recovery/recovery_<model>_<condition>.csv`.

```{r overview-table, results='asis'}
allc <- recovery_all_cors()
if (is.null(allc)) {
  cat("*No recovery runs found yet (`mcmc/recovery/`). ",
      "Run `Rscript modeling/run_recovery.R`.*\n\n")
} else {
  allc %>%
    group_by(model, condition) %>%
    summarise(`Mean r` = sprintf("%.2f", mean(r)),
              `Min r`  = sprintf("%.2f", min(r)),
              n = dplyr::first(n), .groups = "drop") %>%
    kbl(align = c("l", "l", "c", "c", "c"),
        caption = "Recovery correlations summarised per model and condition.") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, position = "center") %>%
    print()
}
```

```{r per-model, results='asis', fig.width=14, fig.height=8}
for (m in c("stl", "fourpar", "ewmv")) {
  lab <- c(stl = "STL", fourpar = "4PAR", ewmv = "EWMV")[[m]]
  cat("\n\n# ", lab, "\n\n", sep = "")
  g <- recovery_grid(m, base_size = 13)
  if (is.null(g)) {
    cat("*Not run yet — `Rscript modeling/run_recovery.R ", m, "`.*\n\n", sep = "")
    next
  }
  print(g$grid)
  cat("\n\n")
  print(
    g$table %>%
      kbl(align = c("l", rep("c", ncol(g$table) - 1)),
          caption = paste0(lab, " parameter recovery correlations.")) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                    full_width = FALSE, position = "center")
  )
  cat("\n\n")
}
```
