---
title: "EWMV Parameter Recovery & Simulated Data"
author: "Zach Tefertiller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(kableExtra); library(here)
})
source(here("modeling", "ewmv", "simulate_ewmv.R"))
source(here("analysis", "plotting", "ewmv_recovery_plot.R"))
```

# Overview

We assess the EWMV model two ways:

1. **Generative validity** — data simulated from the fitted parameters (on the real
   fixed blue / pre-reversal breakpoints) should reproduce the observed behaviour.
2. **Parameter recovery** — refitting the model to the simulated data should return
   the parameters used to generate it. We do this for **realistic** parameters
   (posterior means from the real fit) and for a **fully synthetic** block whose
   parameters are drawn from the model's group prior.

The simulator (`modeling/ewmv/simulate_ewmv.R`) reproduces the `ewmv_vec.stan`
decision process exactly: a within-trial pumping probability
`logistic(τ · u_pump)` with belief `p_burst` updated between trials by the
EWMV learning rule.

# Generative Validity (simulated vs. observed)

```{r ppc, results='asis'}
data <- read.csv(here("data", "processed", "filtered_data", "full_all_recruitments.csv"))
blue <- data %>% filter(balloon_color == "b", trial_number < 91)

sim_path <- here("mcmc", "ewmv_recovery", "sim_block_fitted.csv")
if (file.exists(sim_path)) {
  sim <- read.csv(sim_path)
  smry <- function(df) df %>% group_by(participant_id) %>%
    summarise(adj = mean(inflations[popped == 0], na.rm = TRUE),
              burst = mean(popped), .groups = "drop")
  ss <- smry(sim); rs <- smry(blue)
  tibble(
    Source = c("Observed (real)", "Simulated (EWMV)"),
    `Mean adjusted pumps` = sprintf("%.2f (SD %.2f)",
        c(mean(rs$adj, na.rm = TRUE), mean(ss$adj, na.rm = TRUE)),
        c(sd(rs$adj, na.rm = TRUE),   sd(ss$adj, na.rm = TRUE))),
    `Burst rate` = sprintf("%.2f", c(mean(rs$burst), mean(ss$burst)))
  ) %>%
    kbl(align = c("l", "c", "c"),
        caption = "EWMV-simulated data vs. observed (blue / pre-reversal).") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, position = "center") %>%
    print()
} else {
  cat("*Simulated block not generated yet ",
      "(`mcmc/ewmv_recovery/sim_block_fitted.csv`).*\n\n")
}
```

# Parameter Recovery — Realistic Parameters

True parameters = posterior means from the real blue (pre-reversal) EWMV fit.
Points are participants; the dashed line is identity, the grey line is the fit.

```{r recovery-fitted, fig.width=14, fig.height=3.6, results='asis'}
rec <- ewmv_recovery_report(here("mcmc", "ewmv_recovery", "recovery_fitted.csv"))
if (is.null(rec)) {
  cat("*Recovery not finished yet (`mcmc/ewmv_recovery/recovery_fitted.csv`).*\n\n")
} else {
  print(rec$plot)
  rec$summary %>%
    kbl(align = c("l", "c", "c"),
        caption = "Parameter recovery correlations (realistic parameters).") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, position = "center") %>%
    print()
}
```

# Totally Simulated Block — Prior-Drawn Parameters

A fully synthetic dataset whose parameters are drawn from the EWMV group prior
(no dependence on the real data), then refit.

```{r recovery-prior, fig.width=14, fig.height=3.6, results='asis'}
recp <- ewmv_recovery_report(here("mcmc", "ewmv_recovery", "recovery_prior.csv"),
                             dot_color = "#7209B7")
if (is.null(recp)) {
  cat("*Synthetic block not finished yet (`mcmc/ewmv_recovery/recovery_prior.csv`).*\n\n")
} else {
  print(recp$plot)
  recp$summary %>%
    kbl(align = c("l", "c", "c"),
        caption = "Parameter recovery correlations (prior-drawn parameters).") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, position = "center") %>%
    print()
}
```
